---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

# camtrapmonitoring


## Installation

You can install the development version of camtrapmonitoring with:

```{r, eval = FALSE}
remotes::install_github("robitalec/camtrapmonitoring")
```


This package depends on `sf`, `terra`, and `distanceto`. 

More details on installation here:

- [`sf`: Installing](https://r-spatial.github.io/sf/#installing)
- [`terra`: Installing](https://rspatial.github.io/terra/#installation)



## Name

This package was previously named {wildcam} but renamed to {camtrapmonitoring} 
to reflect broadened scope of package from planning camera trap surveys to 
community based monitoring of wildlife and density estimation, and to avoid
confusion with other projects using camera traps named wildcam or similar. 
