---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = '#>',
  fig.path = 'man/figures/README-',
  out.width = '100%'
)
```


# camtrapmonitoring

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active) 
[![camtrapmonitoring status badge](https://robitalec.r-universe.dev/badges/camtrapmonitoring)](https://robitalec.r-universe.dev/camtrapmonitoring) 
[![CRAN](https://www.r-pkg.org/badges/version/camtrapmonitoring)](https://cran.r-project.org/package=camtrapmonitoring)

## Installation

You can install the development version of camtrapmonitoring with:

```{r, eval = FALSE}
# r-universe
install.packages('camtrapmonitoring', repos = c('https://robitalec.r-universe.dev', 'https://cloud.r-project.org'))

# Or {remotes}
remotes::install_github('robitalec/camtrapmonitoring')
```


This package depends on {sf}, {terra}, and {distanceto}. 

More details on installation here:

- [`sf`: Installing](https://r-spatial.github.io/sf/#installing)
- [`terra`: Installing](https://rspatial.github.io/terra/#installation)



## Name

This package was previously named {wildcam} but renamed to {camtrapmonitoring} 
to reflect broadened scope of package from planning camera trap surveys to 
community based monitoring of wildlife and density estimation, and to avoid
confusion with other projects using camera traps named wildcam or similar. 
